---
title: "Causes of Death in France and New York"
author: "Charles Martinez (cmm5627), Happy Situ (hys82), Jenna Webb (jnw653)"
date: "October 14, 2015"
output: html_document
---

For this project we found two different data sets--each representing different causes of death in New York and France. We then used various methods to join the two data sets and then displayed interesting plots about them.  

####This is the code that we used to be able to put the data into an Oracle database:  

```
require(tidyr)
require(dplyr)
require(ggplot2)

setwd("C:/Users/jenna_000/Jennas_Stuff/UT/DataVisualization/DV_RProject3/01 Data")

file_path <- "CausesOfDeath_France_2001-2008.csv"

France <- read.csv(file_path, stringsAsFactors = FALSE)

#str(France) # Uncomment this and  run just the lines to here to get column types to use for getting the list of measures.

France_measures <- c("YEAR", "Value")

dimensions <- setdiff(names(France), France_measures)
if( length(France_measures) > 1 || ! is.na(dimensions)) {
  for(d in dimensions) {
    # Get rid of " and ' in dimensions.
    France[d] <- data.frame(lapply(France[d], gsub, pattern="[\"']",replacement= ""))
    # Change & to and in dimensions.
    France[d] <- data.frame(lapply(France[d], gsub, pattern="&",replacement= " and "))
    # Change : to ; in dimensions.
    France[d] <- data.frame(lapply(France[d], gsub, pattern=":",replacement= ";"))
  }
}

library(lubridate)

# The following is an example of dealing with special cases like making state abbreviations be all upper case.
# df["State"] <- data.frame(lapply(df["State"], toupper))

# Get rid of all characters in measures except for numbers, the - sign, and period.dimensions
if( length(France_measures) > 1 || ! is.na(France_measures)) {
  for(m in France_measures) {
    France[m] <- data.frame(lapply(France[m], gsub, pattern="[^--.0-9]",replacement= ""))
  }
}

write.csv(France, paste(gsub(".csv", "", file_path), ".reformatted.csv", sep=""), row.names=FALSE, na = "")

tableName <- gsub(" +", "_", gsub("[^A-z, 0-9, ]", "", gsub(".csv", "", file_path)))
sql <- paste("CREATE TABLE", tableName, "(\n-- Change table_name to the table name you want.\n")
if( length(France_measures) > 1 || ! is.na(dimensions)) {
  for(d in dimensions) {
    sql <- paste(sql, paste(d, "varchar2(4000),\n"))
  }
}
if( length(France_measures) > 1 || ! is.na(France_measures)) {
  for(m in France_measures) {
    if(m != tail(France_measures, n=1)) sql <- paste(sql, paste(m, "number(38,4),\n"))
    else sql <- paste(sql, paste(m, "number(38,4)\n"))
  }
}
sql <- paste(sql, ");")
cat(sql)

file_path2 <- "New_York_City_Leading_Causes_of_Death.csv"

NewYork <- read.csv(file_path2, stringsAsFactors = FALSE)

#str(NewYork)

NewYork_measures <- c("YEAR", "COUNT", "PERCENT")

NY_dimensions <- setdiff(names(NewYork), NewYork_measures)
if( length(NewYork_measures) > 1 || ! is.na(NY_dimensions)) {
  for(d in NY_dimensions) {
    # Get rid of " and ' in dimensions.
    NewYork[d] <- data.frame(lapply(NewYork[d], gsub, pattern="[\"']",replacement= ""))
    # Change & to and in dimensions.
    NewYork[d] <- data.frame(lapply(NewYork[d], gsub, pattern="&",replacement= " and "))
    # Change : to ; in dimensions.
    NewYork[d] <- data.frame(lapply(NewYork[d], gsub, pattern=":",replacement= ";"))
  }
}

# Get rid of all characters in measures except for numbers, the - sign, and period.dimensions
if( length(NewYork_measures) > 1 || ! is.na(NewYork_measures)) {
  for(m in NewYork_measures) {
    NewYork[m] <- data.frame(lapply(NewYork[m], gsub, pattern="[^--.0-9]",replacement= ""))
  }
}

write.csv(NewYork, paste(gsub(".csv", "", file_path2), ".reformatted.csv", sep=""), row.names=FALSE, na = "")

tableName <- gsub(" +", "_", gsub("[^A-z, 0-9, ]", "", gsub(".csv", "", file_path2)))
sql <- paste("CREATE TABLE", tableName, "(\n-- Change table_name to the table name you want.\n")
if( length(NewYork_measures) > 1 || ! is.na(NY_dimensions)) {
  for(d in NY_dimensions) {
    sql <- paste(sql, paste(d, "varchar2(4000),\n"))
  }
}
if( length(NewYork_measures) > 1 || ! is.na(NewYork_measures)) {
  for(m in NewYork_measures) {
    if(m != tail(NewYork_measures, n=1)) sql <- paste(sql, paste(m, "number(38,4),\n"))
    else sql <- paste(sql, paste(m, "number(38,4)\n"))
  }
}
sql <- paste(sql, ");")
cat(sql)
```

####Here is a summary of our data:  
```{r}
source("../01 Data/Death_jnw.R", echo = FALSE)
```

```{r}
summary(df_France)
summary(df_NewYork)
```

####Here is a summary and a subset after an inner join, and the code that was used to join:
```
innerj <- dplyr::inner_join(dfFrance, dfNYC, by="YEAR")
```


####A bind and its code:  

```{r}
source("../01 Data/Death_cmm.R", echo = FALSE)
```

```
df_bind <- dplyr::left_join(df_NewYorkDeaths,df_FranceDeaths,by="YEAR") 
df_bind <- dplyr::distinct(df_bind)
df_bind <- dplyr::rename(df_bind, NY_SEX = SEX.x)
df_bind <- dplyr::rename(df_bind, FRANCE_SEX = SEX.y)
```
```{r}
summary(df_bind)
head(df_bind)
```

####And a full join, as well as its code:  
```
df_full <- dplyr::full_join(df_France, df_NewYork, by="YEAR")
```

```{r}
summary(df_bind)
head(df_bind)
```

```{r}
summary(df_full)
head(df_full)
```
  
###Plots 1 & 2:
```{r}
source("../02 Data Wrangling/InnerJoin.R", echo = TRUE)
```

  
###Plot 3:  
```{r}
source("../02 Data Wrangling/Left_Join_Wrangle.R", echo = TRUE)
```

The plot above compares the number of accidental deaths by sex in New York and France. What is interesting about this graph is that it offers a visualization that provides the viewer with eight possible comparisons. So, one can compare the number of accidental deaths between males in both France and New York as well as the number of accidental deaths between males and females only in New York or in France. The data was chosen by inquiry into which location offered more of a risk of death due to accident. From the data, one could assert that males who live in New york are more likely to die from an accident than males in France. This could be due to the density of the population in urban location. Additionally, one could also posit that females in France and New York, in general, are less likely to die from accidents than males from both France and New York. 

###Plots 4 & 5:  
```{r}
source("../02 Data Wrangling/full_data_wrangle.R", echo = TRUE)
```
  
In the preceding plots, the suicide rates were looked at for both France and New York for the years from 2007-2008. For France, it seems that males tended to have a higher suicide rate tahn females, and it also seems that there were few suicides in the middle of the year. For New York, the deaths tended to be more consistent throughout the year, but again males had a higher suicide rate than females.  
  
